A sheaf--function dictionary for algebraic tori over local fields

### 2013-08-20 shanghai.tex
This version has a proposition in the last section, but I think the proof sketched there is not correct.

We will also need the Greenberg functor, 
introduced in \cite{greenberg:61} and \cite{greenberg:63a}, 
as a functor from the category of schemes locally of finite type 
over an Artinian local ring with perfect residue field $k$ 
to the category of schemes locally of finite type over $k$. 
This functor respects open immersions, closed immersions, 
etale morphisms, smooth morphisms and geometric components.
Besides Greenberg's original papers, see [DG70, V, § 4, no. 1])
see \cite{bosch-lutkebohmert-reynaud:NeronModels}*{Ch. 9, \S 6} 
and the first few sections of \cite{stasinski:12a} 
for background on the Greenberg functor.

For any positive integer $n$, let
\[
%\begin{tikzcd}
\Gr_n : \left( \text{Schemes} \atop \text{lft over $\OK$}\right) \to \left( \text{Schemes} \atop \text{lft over $\Fq$}\right)
%\end{tikzcd}
\]
be the functor produced by precomposing pull-back along $\OK \to \OK/\pK^n$ with the Greenberg functor.
The surjective ring homomorphism $\OK/\pK^{n+1} \to \OK/\pK^n$ determines a natural transformation $P_{n} : \Gr_{n+1} \to \Gr_n$ 
%\Clifton{Do you think I should say more about $P_{n}$?}
\cite{bosch-lutkebohmert-reynaud:NeronModels}*{p. 276}.
For a given Néron model $\TT$ of $T$, set
\[
\GN{T}_n \ceq \Gr_n(\TT)
\qquad \text{and}\qquad
f_n \ceq P_n(\TT).
\]
Some fundamental properties of $f_n : \GN{T}_{n+1} \to \GN{T}_n$ are:
 \begin{enumerate}
   \labitem{(GN.0)}{points} $\GN{T}_n(\Fq) = \TT(\OK/\pK^n)$, canonically.
  \labitem{(GN.1)}{smooth} 
  $\GN{T}_n$ is a smooth commutative group scheme locally of finite type over $\Fq$ 
  and the component group of $\GN{T}_n$ is geometrically finitely generated. 
  Note that $\underline{T}$ is a smooth group scheme over $\OK$, by construction. 
  Since smoothness is preserved under base change, 
  $\TT_n \ceq \underline{T} \times_\Spec{\OK} \Spec{\OK/\pK^n}$ 
  is a smooth group scheme over an Artinian ring with perfect residue field. 
  The Greenberg functor preserves smoothness, 
  so $\GN{T}_n$ is a smooth group scheme locally of finite type over $\Fq$. 
  The Greenberg functor takes the geometric identity component of $\TT_n$ 
  to the geometric identity component of $\GN{T}_n$;
  in fact, $\pi_0(\GN{T}_n) = \pi_0(\TT) \times_{\Spec{\OK}} \Spec{\Fq}$ and
  this group is finitely generated by \cite{xarles:93a}*{2.18}.
  \labitem{(GN.2)}{truncation}
  The morphism $f_n : \GN{T}_{n+1} \to \GN{T}_n$ is a morphism of group schemes over $\Fq$. 
  This follows from the fact that $P_{n}$ is a natural transformation between additive functors.
  We remark that $f_n : \GN{T}_{n+1} \to \GN{T}_n$ is not induced from a morphism of Néron models.
\end{enumerate}

@@

\begin{proof}
Weil restriction behaves properly with respect to formation of Néron models:
there is a canonical isomorphism
\[
\underline{\Res_{K'/K}(T')} \iso \Res_{\mathcal{O}_{K'}/\OK}(\underline{T'}).
\]
And Weil restriction behaves properly with respect to the Greenberg transform:
there is a canonical isomorphism 
\Clifton{References needed!}
\[
\Gr_{n}\left( \Res_{\mathcal{O}_{K'}/\OK}(\underline{T'})\right)
\iso
\Res_{k'/k} \Gr_n(\underline{T'}),
\]
where $k'$ is the residue field for $K'$.
And the Greenberg tranforms respects products:
\[
\Gr_n(\underline{T'}) = \Gr_n(\underline{T}\times_{k} {k'}) \iso \Gr_n(\underline{T})\times_{k} {k'} = \Gr_n(\underline{T})_{k'}.
\]
Thus, the 
Now use Proposition~\ref{prop:csbe}.
\[
\begin{tikzcd}
\QCiso{\Res_{k'/k}(G_{k'})} \arrow[two heads]{r}{\iota^*} \dar[swap]{\trFrob{\Res_{k'/k}(G_{k'})}} & \QCiso{G} \dar{\trFrob{G}} \\
G(k')^* \arrow[two heads]{r}{i_{k'/k}^*} & G(k)^*.
\end{tikzcd}
\]

\end{proof}

>>>>

We will also need the Greenberg functor, 
introduced in \cite{greenberg:61} and \cite{greenberg:63a}, 
as a functor from the category of schemes locally of finite type 
over an Artinian local ring with perfect residue field $k$ 
to the category of schemes locally of finite type over $k$. 
This functor respects open immersions, closed immersions, 
etale morphisms, smooth morphisms and geometric components.
Besides Greenberg's original papers, see \cite{demazure-gabriel:GroupesAlgebriques}*{V, \S 4, no. 1}, \cite{bosch-lutkebohmert-reynaud:NeronModels}*{Ch. 9, \S 6} 
and the first few sections of \cite{stasinski:12a} 
for background on the Greenberg functor.

For any positive integer $n$, let
\[
%\begin{tikzcd}
\Gr_n : \left( \text{Schemes} \atop \text{lft over $\OK$}\right) \to \left( \text{Schemes} \atop \text{lft over $\Fq$}\right)
%\end{tikzcd}
\]
be the functor produced by precomposing pull-back along $\OK \to \OK/\pK^n$ with the Greenberg functor.
The surjective ring homomorphism $\OK/\pK^{n+1} \to \OK/\pK^n$ determines a natural transformation $P_{n} : \Gr_{n+1} \to \Gr_n$.
%\Clifton{Do you think I should say more about $P_{n}$?}
%\cite{bosch-lutkebohmert-reynaud:NeronModels}*{p. 276}.
\Clifton{I took the reference to BLR for $P_n$ out, because it didn't actually give any information. This must be in DG, but I don't have a copy.}
For a given Néron model $\TT$ of $T$, set
\[
\GN{T}_n \ceq \Gr_n(\TT)
\qquad \text{and}\qquad
f_n \ceq P_n(\TT).
\]
Some fundamental properties of $f_n : \GN{T}_{n+1} \to \GN{T}_n$ are:
 \begin{enumerate}
   \labitem{(GN.0)}{points} $\GN{T}_n(\Fq) = \TT(\OK/\pK^n)$, canonically.
  \labitem{(GN.1)}{smooth} 
  $\GN{T}_n$ is a smooth commutative group scheme locally of finite type over $\Fq$ 
  and the component group of $\GN{T}_n$ is geometrically finitely generated. 
  Note that $\underline{T}$ is a smooth group scheme over $\OK$, by construction. 
  Since smoothness is preserved under base change, 
  $\TT_n \ceq \underline{T} \times_\Spec{\OK} \Spec{\OK/\pK^n}$ 
  is a smooth group scheme over an Artinian ring with perfect residue field. 
  The Greenberg functor preserves smoothness, 
  so $\GN{T}_n$ is a smooth group scheme locally of finite type over $\Fq$. 
  The Greenberg functor takes the geometric identity component of $\TT_n$ 
  to the geometric identity component of $\GN{T}_n$;
  in fact, $\pi_0(\GN{T}_n) = \pi_0(\TT) \times_{\Spec{\OK}} \Spec{\Fq}$ and
  this group is finitely generated by \cite{xarles:93a}*{2.18}.
  \labitem{(GN.2)}{truncation}
  The morphism $f_n : \GN{T}_{n+1} \to \GN{T}_n$ is a morphism of group schemes over $\Fq$. 
  This follows from the fact that $P_{n}$ is a natural transformation between additive functors.
  We remark that $f_n : \GN{T}_{n+1} \to \GN{T}_n$ is not induced from a morphism of Néron models.
\end{enumerate}

@@

\begin{proof}
Weil restriction behaves properly with respect to formation of Néron models:
adapting the argument in \cite{bosch-lutkebohmert-reynaud:NeronModels}*{\S 7.6, Prop 6} 
gives a canonical isomorphism 
\Clifton{I need to check this.}
\begin{equation}
\underline{\Res_{K'/K}(T')} \iso \Res_{\mathcal{O}_{K'}/\OK}(\underline{T'}).
\end{equation}
And Weil restriction behaves properly with respect to the Greenberg transform:
\Clifton{I'm not so sure about this . . . In fact, I don't believe it!}
\begin{equation}
\Gr_{n}\left( \Res_{\mathcal{O}_{K'}/\OK}(\underline{T'})\right)
\iso
\Res_{k'/k} \Gr_n(\underline{T'}),
\end{equation}
where $k'$ is the residue field for $K'$.
And the Greenberg transform respects products:
\[
\Gr_n(\underline{T'}) = \Gr_n(\underline{T}\times_{k} {k'}) \iso \Gr_n(\underline{T})\times_{k} {k'} = \Gr_n(\underline{T})_{k'}.
\]
Thus, the 
Now use Proposition~\ref{prop:csbe}.
\end{proof}
